Abstractions for Type-Level Programming

### 2022-05-10 article.tex
Revert "More diff"
This reverts commit 0cde87495c11c430f41f6e19a3c53dc7d20732d7.

\subsection{Comparison to Generalized Singleton Types}
\begin{diff}
In \Cref{chap:generalizing-singleton-types}, we presented a generalization of singleton types for type-level programing.
Generalized singletons and match types share some similarities, such as the notion of disjointness, which plays an important role in both features.
The two features relate on a deeper level: we can explain match types in terms of generalized singletons.
For example, we can give an alternative definition of the |Elem| match type using a generalized singleton:
%
\begin{lstlisting}
type Elem[X] = { (_: X) match
  case _: String => (_: Char)
  case _: List[t] => (_: Elem[t])
  case _: Any => (_: X) }
\end{lstlisting}
%
This definition makes the term/type correspondence evident: the |Elem[X]| type reduces analogously to the corresponding match expression.
\end{diff}

@@

Scala tuples, for example, fall into this category.
|Tuple2|, the class for pairs, is defined as follows: |case class Tuple2[+T1, +T2](_1: T1, _2: T2)|.
\begin{diff}
Given two instantiations of this this type, |Tuple2[Int,X]| and |Tuple2[String,X]|, although |Tuple2[Nothing,X]| is a subtype of both, there is no runtime value of type |Tuple2[Nothing,X]| (since |Nothing| is uninhabited), and, as a result, those two types are disjoint.
Our disjointness algorithm implements this kind of reasoning to conclude disjointness in the presence of covariant type parameters.
There is no such reasoning available for contravariant types: our algorithm can never conclude that |G[X]| and |G[Y]| are disjoint for a contravariant type |G|.
\end{diff}

>>>>

\subsection{Comparison to Generalized Singleton Types}

In \Cref{chap:generalizing-singleton-types}, we presented a generalization of singleton types for type-level programing.
Generalized singletons and match types share some similarities, such as the notion of disjointness, which plays an important role in both features.
The two features relate on a deeper level: we can explain match types in terms of generalized singletons.
For example, we can give an alternative definition of the |Elem| match type using a generalized singleton:
%
\begin{lstlisting}
type Elem[X] = { (_: X) match
  case _: String => (_: Char)
  case _: List[t] => (_: Elem[t])
  case _: Any => (_: X) }
\end{lstlisting}
%
This definition makes the term/type correspondence evident: the |Elem[X]| type reduces analogously to the corresponding match expression.

@@

Scala tuples, for example, fall into this category.
|Tuple2|, the class for pairs, is defined as follows: |case class Tuple2[+T1, +T2](_1: T1, _2: T2)|.
Given two instantiations of this this type, |Tuple2[Int,X]| and |Tuple2[String,X]|, although |Tuple2[Nothing,X]| is a subtype of both, there is no runtime value of type |Tuple2[Nothing,X]| (since |Nothing| is uninhabited), and, as a result, those two types are disjoint.
Our disjointness algorithm implements this kind of reasoning to conclude disjointness in the presence of covariant type parameters.
\begin{diff}
There is no such reasoning available for contravariant types: our algorithm can never conclude that |G[X]| and |G[Y]| are disjoint for a contravariant type |G|.
\end{diff}

### 2022-05-10 symposium.tex
Revert "More diff"
This reverts commit 0cde87495c11c430f41f6e19a3c53dc7d20732d7.

Stephanie Weirich presented on several occasions a regular expression library as an example of a dependently-typed program written in Haskell \citep{weirich2014examples}.
The library provides safety guarantees similar to ours, but from the slightly different angle of named-capturing groups.
An important conceptual difference between both approaches is ours only focuses on type checking (we rely on Java's regular expressions at run-time), whereas hers also includes a fully-fledged regular expression engine.

>>>>

Stephanie Weirich presented on several occasions a regular expression library as an example of a dependently-typed program written in Haskell \citep{weirich2014examples}.
The library provides safety guarantees similar to ours, but from the slightly different angle of named-capturing groups.
\begin{diff}
An important conceptual difference between the two approaches is that ours only focuses on type checking (we rely on Java's regular expressions at run-time), whereas hers also includes a fully-fledged regular expression engine.
\end{diff}
